---
title: Explore and Plot by Vector Layer Attributes
teaching: 40
exercises: 20
source: Rmd
---

```{r setup, echo=FALSE}
source("setup.R")
```

::::::::::::::::::::::::::::::::::::::: objectives

- Query attributes of a spatial object.
- Subset spatial objects using specific attribute values.
- Plot a vector feature, colored by unique attribute values.

::::::::::::::::::::::::::::::::::::::::::::::::::

:::::::::::::::::::::::::::::::::::::::: questions

- How can I compute on the attributes of a spatial object?

::::::::::::::::::::::::::::::::::::::::::::::::::

```{r load-libraries, echo=FALSE, results="hide", message=FALSE}
library(terra)
library(ggplot2)
library(dplyr)
library(sf)
```

```{r load-data, echo=FALSE, results="hide"}
# learners will have this data loaded from previous episodes
point_harv <-
    st_read("data/NEON-DS-Site-Layout-Files/HARV/HARVtower_UTM18N.shp")
lines_harv <- st_read("data/NEON-DS-Site-Layout-Files/HARV/HARV_roads.shp")
aoi_boundary_harv <-
    st_read("data/NEON-DS-Site-Layout-Files/HARV/HarClip_UTMZ18.shp")
```

::::::::::::::::::::::::::::::::::::::::::  prereq

## Things You'll Need To Complete This Episode

See the [lesson homepage](.) for detailed information about the software, data,
and other prerequisites you will need to work through the examples in this
episode.



::::::::::::::::::::::::::::::::::::::::::::::::::



## Load the Data

We will continue using the `sf`, `terra` `dplyr` and `ggplot2` packages in this
episode. Make sure that you have these packages loaded. We will continue to
work with the three ESRI `shapefiles` (vector layers) that we loaded in the
[Open and Plot Vector Layers in R](https://datacarpentry.org/r-raster-vector-geospatial/06-vector-open-shapefile-in-r) episode.




